---
title: "Homework 3"
subtitle: "Part 3"
output: github_document
author: Ronae McLin rkm2147
---


```{r}
library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d

scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

```{r}
library(p8105.datasets)
data("instacart")
```



- and giving illustrative examples of observations. Then, do or answer the following (commenting on the results of each):



Instacart is a online service that allows for users to shop locally from vairous stores that can then be delivered within NYC. There are a totla of 1,384,617 observations contained within this data.
This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Observations are the level of items in orders by user. Variables include: `r names(instacart)`. 

**There are 134 aisles, and aisles in which the top 3 most ordered items are ordered from are fresh vegetables, fresh fruits, and packaged vegetables & fruits** 

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```


**Plot of items ordered from each aisle, more than 10000.**

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  ggplot(aes( x = aisle, y = n)) + geom_point() + theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1))
```

**A table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.**

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```

**A table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week**

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  knitr::kable()
```


